```{r setup, include=FALSE}
library(tidyverse)
library(igraph)
library(stringr)
library(ggforce)
library(data.table)
library(ggupset)
library(gggenes)
library(forcats)
library(ape)
library(RColorBrewer)
library(svglite)
library(aplot)
library(ggh4x)
library(treeio)
library(ggtree)
library(ggtreeExtra)

`%ni%` <- Negate(`%in%`)
```


## Loading region data for all four genome catalogues

### Considerations for the different databases
Database sizes:
HumGut: 30691 MAGs
GOMC: 24195 MAGs
REThiNk: 53501 MAGs
MFD: 23843 MAGs

Quality:
All databases are a mix of HQ and MQ MAGs (>50% completeness, <10% contamination (<5% contamination for HumGut))


Redundancy:
HumGut is dereplicated to 5170 at 95%
REThiNk is dereplicated to 25222 MAGs at 95% ANI with dRep

Number of MAGs when all is HQ and all is dereplicated:
HumGut: 2094 MAGs
REThiNk: 12106 MAGs


```{r}

### HumGut and GOMC were using an old GTDB version, so we updated that:
GOMC_HumGut_tax <- fread("data/GTDB220_HumGut_GOMC/gtdbtk-v2.4.0_GOMC-HumGut.summary.tsv") %>% rename(genome_id = user_genome, gtdb_classification = classification)

##########
# HumGut #
##########
HumGut_tax <- fread("data/HumGut/HumGut.tsv") %>%
    mutate(genome_file = gsub(".fna.gz", "", genome_file)) %>%
    select(genome_id = genome_file, OLD_gtdb_classification = gtdbtk_taxonomy, gtdbtk_organism_name, ncbi_organism_name, `Contamination (%)` = contamination, 
          `Completeness (%)` = completeness,     cluster95) %>%
    mutate(Database = "HumGut") %>%
    left_join(GOMC_HumGut_tax, by = "genome_id")
  
    


HumGut_all_regions <- fread("data/HumGut/all_regions.tsv") %>%
  rename(genome_id = file) %>%
  separate_longer_delim(product, delim = " / ") %>%
  mutate(
    Category = ifelse(
      str_detect(product,
        "wzy_dependent_putative|wzy_dependent_putative_NO_OPX"
      ), "Uncharacterised", "NA"
    ),
    Category = ifelse(
      str_detect(product,
        "ABC_dependent_putative|ABC_dependent_putative_NO_OPX"
      ), "Uncharacterised", Category
    ),
    Category = ifelse(
      str_detect(product,
        "ABC_wzy_dependent_putative|ABC_wzy_dependent_putative_NO_OPX"
      ), "Uncharacterised", Category
    ),
    Category = ifelse(
      str_detect(product,
      "amylovoran_stewartan|acetan|succinoglycan|colanic|xanthan|galactoglucan|psl|PNAG_eps|lactobacillus_eps|sphingans|zooglan|emulsan|PNAG_eps"
      ), "Wzx/Wzy", Category
    ),
    Category = ifelse(
      str_detect(product, 
      "cellulose|pel-positive|pel|alginate|curdlan|PNAG-ica|PNAG-pga|HA_has"), "Synthase", Category
    ),
    Category = ifelse(
      str_detect(product, 
      "fructansucrase|alpha-glucansucrase"), "Sucrase", Category
    ),
    n_Category = case_when(
      Category == "Uncharacterised" ~ 1,
      Category == "Synthase" ~ 2,
      Category == "Wzx/Wzy" ~ 3,
      Category == "Sucrase" ~ 4
    ),
    product = case_when(
      product == "acetylated-cellulose" ~ "cellulose-acetylated",
      product == "type-1-cellulose" ~ "cellulose-type-1",
      product == "type-2-cellulose" ~ "cellulose-type-2",
      product == "type-3-cellulose" ~ "cellulose-type-3",
      product == "generic-cellulose" ~"cellulose-generic",
      .default = product
    )
    ) %>% left_join(HumGut_tax, by = c("genome_id")) 

HumGut_metadata <- HumGut_all_regions %>% mutate(shared_name = paste0(genome_id, "_", record_id, ".region00", region, ".gbk_region_1")) %>%
  separate(gtdb_classification, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep = ";")

fwrite(HumGut_metadata, "./data/HumGut/HumGut_metadata.tsv", sep = "\t")


###########
# REThiNk #
###########

REThiNk_HQ_tax <- fread("data/REThiNk/22K-hqMAGs.GTDB_midas.taxa.tsv") %>%
  select(genome_id = bin, gtdb_classification = tax_gtdb.r220)

REThiNk_MQ_tax <- fread("data/REThiNk/31K-mqMAGs.GTDB_taxa.tsv") %>%
  select(genome_id = bin, gtdb_classification = tax_gtdb.r220)

REThiNk_all_tax <- rbind(REThiNk_HQ_tax, REThiNk_MQ_tax)  %>%
  mutate(Database = "MiDAS Global")

REThiNk_all_regions <- fread("data/REThiNk/all_regions.tsv") %>%
  rename(genome_id = file) %>%
  separate_longer_delim(product, delim = " / ") %>%
  mutate(
    Category = ifelse(
      str_detect(product,
        "wzy_dependent_putative|wzy_dependent_putative_NO_OPX"
      ), "Uncharacterised", "NA"
    ),
    Category = ifelse(
      str_detect(product,
        "ABC_dependent_putative|ABC_dependent_putative_NO_OPX"
      ), "Uncharacterised", Category
    ),
    Category = ifelse(
      str_detect(product,
        "ABC_wzy_dependent_putative|ABC_wzy_dependent_putative_NO_OPX"
      ), "Uncharacterised", Category
    ),
    Category = ifelse(
      str_detect(product,
      "amylovoran_stewartan|acetan|succinoglycan|colanic|xanthan|galactoglucan|psl|PNAG_eps|lactobacillus_eps|sphingans|zooglan|emulsan|PNAG_eps"
      ), "Wzx/Wzy", Category
    ),
    Category = ifelse(
      str_detect(product, 
      "cellulose|pel-positive|pel|alginate|curdlan|PNAG-ica|PNAG-pga|HA_has"), "Synthase", Category
    ),
    Category = ifelse(
      str_detect(product, 
      "fructansucrase|alpha-glucansucrase"), "Sucrase", Category
    ),
    n_Category = case_when(
      Category == "Uncharacterised" ~ 1,
      Category == "Synthase" ~ 2,
      Category == "Wzx/Wzy" ~ 3,
      Category == "Sucrase" ~ 4
    ),
    product = case_when(
      product == "acetylated-cellulose" ~ "cellulose-acetylated",
      product == "type-1-cellulose" ~ "cellulose-type-1",
      product == "type-2-cellulose" ~ "cellulose-type-2",
      product == "type-3-cellulose" ~ "cellulose-type-3",
      product == "generic-cellulose" ~"cellulose-generic",
      .default = product
    )
    ) %>% left_join(REThiNk_all_tax, by = c("genome_id"))

#######
# All #
#######

l = list(HumGut_all_regions, REThiNk_all_regions)

all_regions <- rbindlist(l, fill = TRUE) %>%
  mutate(
    `Completeness (%)` = ifelse(
    Database %ni% c("HumGut"), 100, `Completeness (%)`
  ),
  `Contamination (%)` = ifelse(
    Database %ni% c("HumGut"), 1, `Contamination (%)`
  )
  )

all_tax <- list(HumGut_tax, REThiNk_all_tax)

all_tax <- rbindlist(all_tax, fill  = TRUE)


```

## Dereplication and HQ MAG subsetting
### HumGut

```{r}

HumGut_all_regions_HQ <- HumGut_all_regions %>% filter(`Completeness (%)` >= 90, `Contamination (%)` <= 5)
HumGut_all_regions_derep <- HumGut_all_regions %>% distinct(cluster95, .keep_all = TRUE)
# Due to how HumGut asks you to filter for the species representatives at 95% ANI (https://github.com/larssnip/HumGut/blob/main/README.md), it matters that you 
#derep before you filter for HQ species if you want only the nes designated as species reps. This results in about 200 less MAGs though.
HumGut_all_regions_derep_HQ <- HumGut_all_regions_derep %>% filter(`Completeness (%)` >= 90, `Contamination (%)` <= 5)

HumGut_tax_derep <- HumGut_tax %>% distinct(cluster95, .keep_all = TRUE) %>% mutate(Database = "HumGut") %>% select(genome_id, gtdb_classification, Database)

# Outputting tsv of the species representative metadata which will be used to select regions for bigscape analysis and for coloring the network in cytoscape:
HumGut_metadata <- HumGut_all_regions_derep %>% 
  mutate(
    region_file_name = paste0(genome_id, "_", record_id, ".region00", region, ".gbk"),
    shared_name = paste0(genome_id, "_", record_id, ".region00", region, ".gbk_region_1")
    ) %>%
  separate(gtdb_classification, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep = ";")

fwrite(HumGut_metadata, "./data/HumGut/HumGut_metadata.tsv", sep = "\t")

HumGut_tax_HQ <- HumGut_tax %>% filter(`Completeness (%)` >= 90, `Contamination (%)` <= 5) %>% mutate(Database = "HumGut") %>% select(genome_id, gtdb_classification, Database)
HumGut_tax_derep_HQ <- HumGut_tax %>% distinct(cluster95, .keep_all = TRUE) %>% filter(`Completeness (%)` >= 90, `Contamination (%)` <= 5) %>% mutate(Database = "HumGut") %>% select(genome_id, gtdb_classification, Database)


HumGut_all_regions <- HumGut_all_regions %>%
  mutate(
    HQ = ifelse(
      genome_id %in% HumGut_all_regions_HQ$genome_id, TRUE, FALSE
    ),
    derep = ifelse(
      genome_id %in% HumGut_all_regions_derep$genome_id, TRUE, FALSE
    ),
    HQ_derep = ifelse(
      genome_id %in% HumGut_all_regions_derep_HQ$genome_id, TRUE, FALSE
    )
  )


p1 <- HumGut_all_regions %>% 
  group_by(Database, Category, product) %>% count() %>% arrange(desc(n)) %>%
  ggplot(aes(x = fct_reorder(product, Category), y = n, fill = product)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Database, ncol = 1) +
  ylab("Number of regions") +
  ggtitle("HumGut All") +
  xlab("") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    legend.position = "none"
    )

p2 <- HumGut_all_regions %>% filter(HQ == TRUE) %>%
  group_by(Database, Category, product) %>% count() %>% arrange(desc(n)) %>%
  ggplot(aes(x = fct_reorder(product, Category), y = n, fill = product)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Database, ncol = 1) +
  ylab("Number of regions") +
  ggtitle("HumGut HQ") +
  xlab("") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    legend.position = "none"
    )

p3 <- HumGut_all_regions %>% filter(derep == TRUE) %>%
  group_by(Database, Category, product) %>% count() %>% arrange(desc(n)) %>%
  ggplot(aes(x = fct_reorder(product, Category), y = n, fill = product)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Database, ncol = 1) +
  ylab("Number of regions") +
  ggtitle("HumGut derep") +
  xlab("") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    legend.position = "none"
    )

p4 <- HumGut_all_regions %>% filter(HQ_derep == TRUE) %>%
  group_by(Database, Category, product) %>% count() %>% arrange(desc(n)) %>%
  ggplot(aes(x = fct_reorder(product, Category), y = n, fill = product)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Database, ncol = 1) +
  ylab("Number of regions") +
  ggtitle("HumGut HQ + derep") +
  xlab("") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    legend.position = "none"
    )

p1 + p2 + p3 + p4

```
### REThiNk

```{r}


REThiNk_derep <- fread("data/REThiNk/derep_MAGs_list.tsv")

REThiNk_all_regions <- REThiNk_all_regions %>%
  mutate(
    HQ = ifelse(
      genome_id %in% REThiNk_HQ_tax$genome_id, TRUE, FALSE
    )
  )

REThiNk_all_regions_derep <- REThiNk_all_regions


REThiNk_all_regions_derep_HQ <- REThiNk_all_regions %>% filter(HQ == TRUE)

REThiNk_tax_HQ <- REThiNk_all_tax %>% filter(genome_id %in% REThiNk_HQ_tax$genome_id) %>% mutate(Database = "MiDAS Global") %>% select(genome_id, gtdb_classification, Database)

REThiNk_tax_derep <- REThiNk_all_tax %>% filter(genome_id %in% REThiNk_derep$genome_id) %>% mutate(Database = "MiDAS Global") %>% select(genome_id, gtdb_classification, Database)

REThiNk_tax_derep_HQ <- REThiNk_all_tax %>% filter(genome_id %in% REThiNk_derep$genome_id & genome_id %in% REThiNk_HQ_tax$genome_id) %>% mutate(Database = "MiDAS Global") %>% select(genome_id, gtdb_classification, Database)


# Outputting tsv of the species representative metadata which will be used to select regions for bigscape analysis and for coloring the network in cytoscape:
REThiNk_metadata <- REThiNk_all_regions_derep %>% 
  mutate(
    region_file_name = paste0(record_id, ".region00", region, ".gbk"),
    shared_name = paste0(genome_id, "_", record_id, ".region00", region, ".gbk_region_1"),
    product = ifelse(Category == "Uncharacterised", "Uncharacterised", product)
    ) %>%
  separate(gtdb_classification, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep = ";")

fwrite(REThiNk_metadata, "./data/REThiNk/REThiNk_metadata.tsv", sep = "\t")

p1 <- REThiNk_all_regions %>% filter(Category != "Uncharacterised") %>%
  group_by(Database, Category, product) %>% count() %>% arrange(desc(n)) %>%
  ggplot(aes(x = fct_reorder(product, Category), y = n, fill = product)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Database, ncol = 1) +
  ylab("Number of regions") +
  ggtitle("REThiNk derep") +
  xlab("") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 15),
    legend.position = "none"
    )


p2 <- REThiNk_all_regions %>% filter(HQ == TRUE)%>% filter(Category != "Uncharacterised") %>%
  group_by(Database, Category, product) %>% count() %>% arrange(desc(n)) %>%
  ggplot(aes(x = fct_reorder(product, Category), y = n, fill = product)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Database, ncol = 1) +
  ylab("Number of regions") +
  ggtitle("REThiNk derep+HQ") +
  xlab("") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 15),
    legend.position = "none"
    )


p1 + p2

```



## Combining the most filtered databases (HQ and derep'ed)

```{r}

l = list(HumGut_all_regions_derep_HQ, REThiNk_all_regions_derep_HQ)

all_regions_HQ_derep <- rbindlist(l, fill = TRUE) %>%
  mutate(
    `Completeness (%)` = ifelse(
    Database %ni% c("HumGut"), 100, `Completeness (%)`
  ),
  `Contamination (%)` = ifelse(
    Database %ni% c("HumGut"), 1, `Contamination (%)`
  ),
  n_Category = case_when(
    Category == "Uncharacterised" ~ 1,
    Category == "Synthase" ~ 2,
    Category == "Wzx/Wzy" ~ 3,
    Category == "Sucrase" ~ 4
  ),
  product = case_when(
    product == "acetylated-cellulose" ~ "cellulose-acetylated",
    product == "type-1-cellulose" ~ "cellulose-type-1",
    product == "type-2-cellulose" ~ "cellulose-type-2",
    product == "type-3-cellulose" ~ "cellulose-type-3",
    product == "generic-cellulose" ~"cellulose-generic",
    .default = product
  )
  )

## Also creating the filtered/derep'ed taxonomy df

l = list(HumGut_tax_derep_HQ, REThiNk_tax_derep_HQ)

all_tax_HQ_derep <- rbindlist(l, fill = TRUE) 

all_tax_regions_HQ_derep <- all_tax_HQ_derep %>% left_join(all_regions_HQ_derep %>% select(-gtdb_classification, -Database), by = "genome_id")

### Outputting HQ and derep'd HumGut genome ids for GTDB-Tk in order to get them up to date (GTDB r220 like REThiNK and MFD)

fwrite(as.list(HumGut_tax_derep_HQ$genome_id), "~/projects/epsSMASH/epsSMASH_paper/data/GTDB220_HumGut_GOMC/paths/HumGut_tax_derep_HQ_ids.txt", sep = "\n")



```

```{r}


l = list(HumGut_all_regions_derep, REThiNk_all_regions_derep)


all_regions_derep <- rbindlist(l, fill = TRUE) %>%
  mutate(
    `Completeness (%)` = ifelse(
    Database %ni% c("GOMC", "HumGut"), 100, `Completeness (%)`
  ),
  `Contamination (%)` = ifelse(
    Database %ni% c("GOMC", "HumGut"), 1, `Contamination (%)`
  ),
  n_Category = case_when(
    Category == "Uncharacterised" ~ 1,
    Category == "Synthase" ~ 2,
    Category == "Wzx/Wzy" ~ 3,
    Category == "Sucrase" ~ 4
  ),
  product = case_when(
    product == "acetylated-cellulose" ~ "cellulose-acetylated",
    product == "type-1-cellulose" ~ "cellulose-type-1",
    product == "type-2-cellulose" ~ "cellulose-type-2",
    product == "type-3-cellulose" ~ "cellulose-type-3",
    product == "generic-cellulose" ~"cellulose-generic",
    .default = product
  )
  )

## Also creating the filtered/derep'ed taxonomy df

l = list(HumGut_tax_derep, REThiNk_tax_derep)

all_tax_derep <- rbindlist(l, fill = TRUE) 

all_tax_regions_derep <- all_tax_derep %>% left_join(all_regions_derep %>% select(-gtdb_classification, -Database), by = "genome_id")

all_tax_regions <- all_tax %>% left_join(all_regions %>% select(-gtdb_classification, -Database), by = "genome_id")

```

## Stats of interest: 
Average/max/min amount of exoPS BGCs?
How many BGCs detected in each database as a % of all genomes?

```{r}

all_tax_regions_derep <- all_tax_regions_derep %>% group_by(genome_id) %>% mutate(
  region_per_genome = ifelse(is.na(product), 0, length(product))
) %>% ungroup()

all_tax_regions_derep %>% ggplot(aes(x = region_per_genome)) +
  geom_bar() +
  facet_wrap(~Database)

all_tax_regions_derep %>% ggplot(aes(x = region_per_genome)) +
  geom_histogram()

# How many percent of genomes do we detect any BGCs in?

all_tax_regions_derep %>%
  group_by(Database) %>% 
  mutate(no_genomes = length(unique(genome_id))) %>% 
  ungroup() %>%
  filter(region_per_genome >= 1) %>%
  group_by(Database) %>%
  mutate(
    no_genomes_w_region = length(unique(genome_id)),
    percent_genomes_w_region = no_genomes_w_region/no_genomes*100
  ) %>% ungroup() %>% select(Database, no_genomes, no_genomes_w_region, percent_genomes_w_region) %>%
  distinct(Database, .keep_all = T)
  
```


## exoPS BGC Category distributions

```{r}

Category_distribution <- all_tax_regions_derep %>% 
  group_by(Database) %>% 
  mutate(no_genomes = length(unique(genome_id))) %>% 
  ungroup() %>%
  filter(!is.na(product), Database == "MiDAS Global") %>%
  group_by(Database, no_genomes, Category, genome_id) %>%
  summarise(product = length(product)) %>%
  group_by(Database, no_genomes, Category) %>% count() %>%
  mutate(`% Genomes with EPS BGC` = (n/no_genomes)*100) %>% arrange(desc(`% Genomes with EPS BGC`)) %>%
  ggplot(aes(x = reorder(Database, n, decreasing = FALSE), y = `% Genomes with EPS BGC`, fill = reorder(Category,n))) +
  geom_bar(stat = "identity") +
  scale_y_continuous(breaks = c(0, 25, 50, 75, 100), limits = c(0,60)) +
  ylab("% Genomes with EPS BGC") +
  xlab("") +
  scale_fill_manual(values = c("#F8766D", "#7CAE00", "#00BFC4", "#C77CFF")) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 20),
    axis.text.y = element_text(size = 20),
    legend.text = element_text(size = 20),
    legend.title = element_blank(),
    axis.title.y = element_text(size = 20)
    )


Category_distribution

ggsave("figures/all/Category_distribution.png", Category_distribution, width = 10, height = 10)

ggsave("figures/all/Category_distribution_no_loose.png", Category_distribution_no_loose, width = 10, height = 10)




```



## exoPS BGC products distributions

```{r}

all_regions_derep_barplot_no_loose <- all_tax_regions_derep %>% 
  group_by(Database) %>% 
  mutate(no_genomes = length(unique(genome_id))) %>% 
  ungroup() %>%
  filter(!is.na(product)) %>%
  filter(Category != "Uncharacterised") %>%
  group_by(genome_id) %>%
  distinct(product, .keep_all = TRUE) %>%
  group_by(Database, no_genomes, n_Category, Category, product) %>% count() %>% arrange(desc(n)) %>%
  mutate(`% Genomes with EPS BGC` = (n/no_genomes)*100) %>%
  ggplot(aes(x = fct_reorder(product, n_Category), y = `% Genomes with EPS BGC`, fill = product)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label= round(`% Genomes with EPS BGC`, digits = 2), vjust = "inward", size = 25)) +
  facet_nested(Database ~ fct_reorder(Category, n_Category), nest_line = element_line(colour = "red"), space = "free", scales = "free_x") +
  scale_y_continuous(trans = "identity") +
  scale_color_brewer(palette = "Set2") +
  xlab("") +
  ylab("% Genomes with EPS BGC") +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1, size = 20),
    axis.text.y = element_text(size = 20, vjust = 0.5),
    strip.text = element_text(size = 25),
    legend.position = "none",
    axis.title.y = element_text(size = 25)
    )
all_regions_derep_barplot_no_loose



ggsave("figures/all/all_regions_derep_barplot_no_loose.png", all_regions_derep_barplot_no_loose, height = 16, width = 24, dpi = 200)




```



### HumGut: Looking into the most prevalent genera in the human gut: 
Bacillota: Streptococcus, Blautia, Faecalibacterium, Lachnoclostridium, Anaerostipes, Ruminococcus
Bacteroidota: Bacteroides
Actinomycetota: Bifidobacterium
Pseudomonadota: Shigella (Escherichia coli)

First off - we see that we lose a lot of information when we only use dereplicated + HQ MAGs. Specifically E. coli/Shigella we completely lose all hits. I think it makes sense to keep the species reps and MQ MAGs in this analysis because we are looking at the most prevalent genera in the human gut, so we want to get the whole picture. 

We see that a good portion of the genomes in all the gram-positive genera have pel gene clusters. Some of the e coli genomes have PNAG and/or cellulose clusters. Half of the Streptococcus genomes have lactobacillus EPS gene clusters, almost half of them contain alpha-glucansucrase genes and a quarter of them contain fructan-sucrase genes. If you look at other genera, there are almost none of these clusters, so this is almost unique to the Streptococcus genus in this environment. 

```{r}


HumGut_tax_Genome_no <- HumGut_tax_derep %>%
  mutate(
    Genus = case_when(
      str_detect(gtdb_classification, "g__Blautia") ~  "Blautia",
      str_detect(gtdb_classification, "g__Streptococcus") ~  "Streptococcus",
      str_detect(gtdb_classification, "g__Faecalibacterium") ~  "Faecalibacterium",
      str_detect(gtdb_classification, "g__Lachnoclostridium|g__Clostridium") ~  "(Lachno-)Clostridium",
      str_detect(gtdb_classification, "g__Anaerostipes") ~  "Anaerostipes",
      str_detect(gtdb_classification, "g__Ruminococcus") ~  "Ruminococcus",
      str_detect(gtdb_classification, "g__Bacteroides") ~  "Bacteroides",
      str_detect(gtdb_classification, "g__Bifidobacterium") ~  "Bifidobacterium",
      str_detect(gtdb_classification, "g__Escherichia") ~  "Escherichia",
      .default = "Other Genera"
    ) 
  ) %>%
  group_by(Genus) %>%
  mutate(Genus_no = sum(length(Genus))) %>%
  select(Genus, Genus_no) %>% distinct(Genus, .keep_all = TRUE)

HumGut_all_regions_prevalent <- all_regions_derep %>% 
  filter(Database == "HumGut", !str_detect(gtdb_classification, "pectinophilus")) %>%
  separate(gtdb_classification, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep = ";") %>%
  mutate(
    Phylum = gsub(".__", "", Phylum),
    Phylum = ifelse(Phylum == "Bacillota_A", "Bacillota", Phylum),
    Genus = case_when(
      str_detect(Genus, "g__Blautia") ~  "Blautia",
      str_detect(Genus, "g__Streptococcus") ~  "Streptococcus",
      str_detect(Genus, "g__Faecalibacterium") ~  "Faecalibacterium",
      str_detect(Genus, "g__Lachnoclostridium|g__Clostridium") ~  "(Lachno-)Clostridium",
      str_detect(Genus, "g__Anaerostipes") ~  "Anaerostipes",
      str_detect(Genus, "g__Ruminococcus") ~  "Ruminococcus",
      str_detect(Genus, "g__Bacteroides") ~  "Bacteroides",
      str_detect(Genus, "g__Bifidobacterium") ~  "Bifidobacterium",
      str_detect(Genus, "g__Escherichia") ~  "Escherichia",
      .default = "Other Genera"
    ) 
  ) %>%
  left_join(HumGut_tax_Genome_no) %>%
  select(genome_id, Phylum, Order, Genus, Genus_no, product, Category, n_Category) %>%
  mutate(
    product = ifelse(Category == "Uncharacterised", "Uncharacterised", product),
    Category = ifelse(Category == "Uncharacterised", "", Category)
         ) %>%
  group_by(genome_id) %>%
  distinct(product, .keep_all = TRUE) %>% 
  ungroup() %>%
  group_by(Genus, n_Category, product) %>%
  mutate(Hits = sum(length(product))) %>% ungroup() %>%
  group_by(Genus) %>%
  distinct(product, .keep_all = TRUE) %>%
  mutate(
    `% genomes with BGC` = round((Hits/Genus_no)*100, digits = 2),
    Genus = ifelse(Genus != "Other Genera", paste0(Phylum, " ", Genus, " ( ", as.character(Genus_no), " )"), paste0(Genus, " ( ", as.character(Genus_no), " )")),
    Order = case_when(
      Order == "o__Actinomycetales" & Genus != "Other Genera ( 25934 )" ~ 1,
      Order == "o__Lactobacillales" & Genus != "Other Genera ( 25934 )" ~ 2,
      Order == "o__Lachnospirales" & Genus != "Other Genera ( 25934 )" ~ 3,
      Order == "o__Oscillospirales" & Genus != "Other Genera ( 25934 )" ~ 4,
      Order == "o__Clostridiales" & Genus != "Other Genera ( 25934 )" ~ 5,
      Order == "o__Bacteroidales" & Genus != "Other Genera ( 25934 )" ~ 6,
      Order == "o__Enterobacterales" & Genus != "Other Genera ( 25934 )" ~ 7,
      .default = 8
    )
    )



HumGut_Heatmap <- ggplot(HumGut_all_regions_prevalent, aes(fct_reorder(product, n_Category), fct_reorder(Genus, Order, .desc = T), fill = `% genomes with BGC`)) +
  geom_tile(aes(), color = "black") +
  scale_fill_gradientn(colors =c("white", "orange", "red"), trans = "identity", limits = c(0,100)) +
  geom_text(aes(label = `% genomes with BGC`), size = 6, color = "black") +
  facet_nested(~ fct_reorder(Category, n_Category), nest_line = element_line(colour = "red"), space = "free", scales = "free_x") +
  xlab("") +
  ylab("") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 25),
    axis.text.y = element_text(size = 25),
    strip.text = element_text(size = 25),
    legend.text = element_text(size = 25),
    legend.title = element_text(size = 25, vjust = 3),
    legend.key.size = unit(2, "cm")
    )
ggsave("figures/humgut/HumGut_Prevalent_Heatmap.png", HumGut_Heatmap, width = 32, height = 16, dpi = 200)

HumGut_all_regions_prevalent_accessions <- all_regions_derep %>% 
  filter(Database == "HumGut", !str_detect(gtdb_classification, "pectinophilus")) %>%
  separate(gtdb_classification, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep = ";") %>%
  filter(
    str_detect(Genus, "g__Blautia|g__Streptococcus|g__Faecalibacterium|g__Lachnoclostridium|g__Clostridium|g__Anaerostipes|g__Ruminococcus|g__Bacteroides|g__Bifidobacterium|g__Escherichia")
    ) %>%
  mutate(shared_name = paste0(genome_id, "_", record_id, ".region00", region, ".gbk_region_1")) %>%
  mutate(record_id = paste0(genome_id, "_", record_id, ".region00", region)) %>%
  mutate(product = ifelse(str_detect(product, "putative"), "Uncharacterised", product))

fwrite(HumGut_all_regions_prevalent_accessions, "data/HumGut/HumGut_prevalent_hits.tsv", sep = "\t")

```

### REThiNk

```{r}
MiDAS_strict <- fread("data/MiDAS5.3_genus_core.txt", sep = "\t") %>% 
  filter(Best == "Strict core") 

REThiNk_tax_Genome_no <- REThiNk_tax_derep %>%
  separate(gtdb_classification, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep = ";") %>%
    mutate(
    Genus = ifelse(
      str_detect(Genus, paste(MiDAS_strict$Genus, collapse = "|")), Genus, "Other Genera"
    ),
    Genus = gsub(".__", "", Genus),
    Phylum = gsub(".__", "", Phylum)
    ) %>%
  group_by(Genus) %>%
  mutate(Genus_no = sum(length(Genus))) %>%
  select(Genus, Genus_no) %>% distinct(Genus, .keep_all = TRUE)
  

REThiNk_all_regions_prevalent <- REThiNk_all_regions_derep %>% 
  filter(gtdb_classification != "Unclassified Bacteria") %>%
  separate(gtdb_classification, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep = ";[a-z]__") %>%
    mutate(
    Genus = ifelse(
      str_detect(Genus, paste(MiDAS_strict$Genus, collapse = "|")), Genus, "Other Genera"
    ),
    Genus = gsub(".__", "", Genus),
    Phylum = gsub(".__", "", Phylum)
    ) %>%
  left_join(REThiNk_tax_Genome_no) %>%
  select(genome_id, Phylum, Family, Genus, Genus_no, product, Category, n_Category) %>%
  mutate(
    product = ifelse(Category == "Uncharacterised", "Uncharacterised", product),
    Category = ifelse(Category == "Uncharacterised", "", Category)
         ) %>%
  group_by(genome_id) %>%
  distinct(product, .keep_all = TRUE) %>% 
  ungroup() %>%
  group_by(Genus, n_Category, product) %>%
  mutate(Hits = sum(length(product))) %>% ungroup() %>%
  group_by(Genus) %>%
  distinct(product, .keep_all = TRUE) %>%
  mutate(
    `% genomes with BGC` = round((Hits/Genus_no)*100, digits = 2),
    label = ifelse(
      Genus != "Other Genera", 
      paste0(Phylum, " ", Genus, " (", as.character(Genus_no), ")"), 
      paste0(Genus, " ( ", as.character(Genus_no), " )")
      ),
    Family = case_when(
      Family == "Flavobacteriaceae" & Genus != "Other Genera ( 21440 )" ~ 1,
      Family == "Chitinophagaceae" & Genus != "Other Genera ( 21440 )" ~ 2,
      Family == "Pirellulaceae" & Genus != "Other Genera ( 21440 )" ~ 3,
      Family == "Sphingomonadaceae" & Genus != "Other Genera ( 21440 )" ~ 4,
      Family == "Clostridiales" & Genus == "Hyphomicrobium_B" ~ 5,
      Family == "Clostridiales" & Genus == "Hyphomicrobium_C" ~ 6,
      Family == "Hyphomicrobiaceae" & Genus != "Other Genera ( 21440 )" ~ 7,
      Family == "Rhodocyclaceae" & Genus != "Other Genera ( 21440 )" ~ 8,
      Family == "Burkholderiaceae" & Genus != "Other Genera ( 21440 )" ~ 9,
      .default = 10
    )
    )


REThiNk_core_heatmap <- ggplot(REThiNk_all_regions_prevalent, aes(fct_reorder(product, n_Category), fct_reorder(label, Family, .desc = T), fill = `% genomes with BGC`), color = "black") +
  geom_tile(aes(), color = "black") +
  ylab("") +
  scale_fill_gradientn(colors =c("white", "orange", "red"), trans = "identity", limits = c(0, 100)) +
  geom_text(aes(label = `% genomes with BGC`), size = 6, color = "black") +
  facet_nested(~ fct_reorder(Category, n_Category), nest_line = element_line(colour = "red"), space = "free", scales = "free_x") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 25),
    axis.text.y = element_text(size = 25),
    strip.text = element_text(size = 25),
    legend.text = element_text(size = 25),
    legend.title = element_text(size = 25, vjust = 3),
    legend.key.size = unit(2, "cm")
    )

ggsave("figures/REThiNk/REThiNk_core_heatmap.png", REThiNk_core_heatmap, width = 32, height = 16, dpi = 200)

REThiNk_strict_accessions <- REThiNk_all_regions_derep %>% 
  filter(gtdb_classification != "Unclassified Bacteria") %>%
  separate(gtdb_classification, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep = ";[a-z]__") %>%
  filter(str_detect(Genus, paste(MiDAS_strict$Genus, collapse = "|"))) %>%
  mutate(file_name = paste0(genome_id, "_", record_id, ".region00", region)) %>%
  mutate(product = ifelse(str_detect(product, "putative"), "Uncharacterised", product))

fwrite(REThiNk_strict_accessions, "data/REThiNk/MiDAS_strict_hits.tsv", sep = "\t")

```


```{r}

taxonomy <- REThiNk_all_regions_derep %>% 
  separate(gtdb_classification, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep = ";") %>%
    mutate(
    Genus = ifelse(
      str_detect(Genus, paste(MiDAS_strict$Genus, collapse = "|")), Genus, "Other Genera"
    ),
    Genus = gsub(".__", "", Genus),
    Phylum = gsub(".__", "", Phylum)
    ) %>%
  select(genome_id, Phylum, Genus, product, Category, n_Category) %>%
  mutate(product = ifelse(Category == "Uncharacterised", "Uncharacterised", product)) %>%
  group_by(Genus, n_Category, product) %>%
  mutate(Hits = sum(length(product))) %>% ungroup() %>%
  left_join(REThiNk_tax_Genome_no) %>%
  filter(genome_id %in% REThiNk_derep$genome_id, genome_id != "089-SG-01.bin.1.675") %>% 
  mutate(
    `% genomes with BGC` = round((Hits/Genus_no)*100, digits = 2),
    tax = ifelse(Genus != "Other Genera", paste0(Phylum, " ", Genus, " ( ", as.character(Genus_no), " )"), paste0(Genus, " ( ", as.character(Genus_no), " )"))
    ) %>% distinct(Genus, .keep_all = TRUE) %>% select(label = genome_id, tax)

REThiNk_tree <- read.tree("../epsSMASH_analyses/REThiNk_catalogue/data/gtdbtk.bac120.decorated.tree")

REThiNk_tree_tibble <- as_tibble(REThiNk_tree) %>% mutate(label = gsub("'", "", label)) %>% left_join(taxonomy) %>% select(-label, label = tax)

REThiNk_tree <- as.phylo(REThiNk_tree_tibble)  %>% keep.tip(tip = taxonomy$tax)

REThiNk_tree <- groupOTU(REThiNk_tree,  split(taxonomy$label, taxonomy$tax))



tree <- ggtree(REThiNk_tree, layout = "rectangular", branch.length = "none") + xlim_tree(0.1)
REThiNk_heatmap <- ggplot(REThiNk_all_regions_prevalent, aes(fct_reorder(product, n_Category), label, fill = `% genomes with BGC`), color = "black") +
  geom_tile(aes(), color = "black") +
  ylab("") +
  scale_fill_gradientn(colors =c("white", "orange", "red"), trans = "identity", limits = c(0, 100)) +
  geom_text(aes(label = `% genomes with BGC`), color = "black") +
  facet_nested(~ fct_reorder(Category, n_Category), nest_line = element_line(colour = "red"), space = "free", scales = "free_x") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 15),
    axis.text.y = element_text(size = 15)
    )

REThiNK_heatmap_tree <- REThiNk_heatmap %>% insert_left(tree, width = 0.1)

ggsave("figures/REThiNk/REThiNk_core_heatmap_tree.png", REThiNK_heatmap_tree, width = 24, height = 16, dpi = 200)

```
