```{r setup, include=FALSE}
library(tidyverse)
library(igraph)
library(stringr)
library(ggforce)
library(data.table)
library(ggupset)
library(gggenes)
library(forcats)
library(readxl)
library(ggrepel)
library(ggfittext)
`%ni%` <- Negate(`%in%`)
```

I want to compare the Synthase-prediction capabilities of epsSMASH to the Bundalovic-Torma paper to see if we miss anything with our method. I've run epsSMASH on the same 1733 genomes they use in their paper (https://doi.org/10.1371/journal.pcbi.1007721) and now I'll show how much they both hit, how much only we hit and how much only they hit. Then I'll investigate what they hit that we dont and vice-versa

```{r}

epsSMASH_all_regions <- fread("data/Bundalovic-Torma/all_regions.tsv") %>%
  filter(str_detect(product, 
      "cellulose|pel-positive|pel|alginate|PNAG-ica|PNAG-pga")) %>% select(Sp_ID = file, epsSMASH = product) %>%
  group_by(Sp_ID) %>%
  mutate(
    epsSMASH = case_when(
      str_detect(epsSMASH, "acetylated-cellulose") ~ "Acetylated-Cellulose",
      str_detect(epsSMASH, "cellulose") ~ "Cellulose",
      str_detect(epsSMASH, "pel") ~ "Pel",
      str_detect(epsSMASH, "alginate") ~ "Alginate",
      str_detect(epsSMASH, "PNAG") ~"PNAG"
    ),
    epsSMASH = paste(epsSMASH, collapse = ", ")
  ) %>%
  select(Sp_ID, epsSMASH) %>%
  distinct(Sp_ID, .keep_all = TRUE)

# Read supplementary file from Bundalovic study (must be added to directory)
Bundalovic_tax <- read_excel("data/Bundalovic-Torma/pcbi.1007721.s015.xlsx", sheet = 1, skip = 1) %>% select(Sp_ID = `NCBI RefSeq Genome ID`, Sp_name = `Species Name`)

# Read supplementary file from Bundalovic study (must be added to directory)
Bundalovic_regions <- read_excel("data/Bundalovic-Torma/pcbi.1007721.s009.xlsx", sheet = 1, skip = 1) %>% 
  unite(col = "operon",7:19) %>%
  select(Sp_name, Sp_ID, `Biofilm Type`, Num_Loci, operon) %>%
  group_by(Sp_ID) %>%
  mutate(
    Bundalovic = paste(`Biofilm Type`, collapse = ", "),
    operon = paste(operon, collapse = " AND "),
    Num_Loci = sum(Num_Loci)
    ) %>%
  ungroup() %>%
  select(Sp_ID, Bundalovic, Num_Loci, operon) %>% 
  distinct(Sp_ID, .keep_all = TRUE)

Bundalovic_regions <- Bundalovic_tax %>% left_join(Bundalovic_regions)


compare <- Bundalovic_regions %>% left_join(epsSMASH_all_regions) %>%
  mutate(
    hits = case_when(
      is.na(Bundalovic) & !is.na(epsSMASH) ~ "epsSMASH",
      !is.na(Bundalovic) & is.na(epsSMASH) ~ "Bundalovic",
      !is.na(Bundalovic) & !is.na(epsSMASH) ~ "Both",
      is.na(Bundalovic) & is.na(epsSMASH) ~ "none"
    )
  ) %>% 
  select(Sp_ID, Bundalovic, epsSMASH, hits, Num_Loci, operon) %>%
  left_join(Bundalovic_tax)


n_compare <- count(group_by(compare, hits), hits) %>% filter(hits != "none")

# Compute percentages
n_compare$fraction = n_compare$n / sum(n_compare$n)

# Compute the cumulative percentages (top of each rectangle)
n_compare$ymax = cumsum(n_compare$fraction)

# Compute the bottom of each rectangle
n_compare$ymin = c(0, head(n_compare$ymax, n=-1))


# Compute label position
n_compare$labelPosition <- (n_compare$ymax + n_compare$ymin) / 2

# Compute a good label
n_compare$label <- paste0(n_compare$hits, "\n Hits: ", n_compare$n)

# Make the plot
ggplot(n_compare, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=hits)) +
  geom_rect() +
  geom_label( x=3.5, aes(y=labelPosition, label=label), size=6) +
  scale_fill_brewer(palette=4) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "none")



```
653 genomes contain synthase-dependent (Pel, PNAG, alginate, cellulose) according to both methods. epsSMASH detects synthase-dependent systems in an additional 10, Bundalovic detects in an additional 96. However the additional hits detected by Bundalovic are often only 2-gene "clusters, or clusters with many duplicates of one gene if there are more than two. Some of them are not detected by the relaxed rules from epsSMASH due to the nature of the rule (e.g. we detect pel-like clusters if we hit pelF and pelG, and some of the Bundalovic pel clusters we dont hit consist of just pelA and pelF, which we did not find to be a good clustering rule generally). Some would have been hit by a relaxed epsSMASH rule, so the reason they are not is either because our HMMs differ or because our bitscore cut-off was more strict than their e-value cut-off at 10⁻5.


# Comparing hits for each synthase system
```{r}


Bundalovic_regions <- read_excel("data/Bundalovic-Torma/pcbi.1007721.s009.xlsx", sheet = 1, skip = 1) %>% 
  unite(col = "operon",7:19) %>%
  select(Sp_ID, Bundalovic=`Biofilm Type`) %>% 
  mutate(Bundalovic = ifelse(Bundalovic=="Acetylated-Cellulose", "Acetylated cellulose", Bundalovic))%>%
  mutate(compare = paste0(Sp_ID, Bundalovic)) %>% select(compare,Bundalovic)

epsSMASH_Bundalovic_all_regions <- fread("data/Bundalovic-Torma//all_regions.tsv") %>%
  filter(str_detect(product, 
      "cellulose|pel-positive|pel|alginate|PNAG-ica|PNAG-pga")) %>% 
  filter(product %ni% c("cyanocellulose", "cellulose-gram-positive", "cellulose-gram-positive-like")) %>%
  select(Sp_ID = file, epsSMASH = product) %>%
  mutate(
    epsSMASH = case_when(
      str_detect(epsSMASH, "acetylated-cellulose") ~ "Acetylated cellulose",
      str_detect(epsSMASH, "cellulose") ~ "Cellulose",
      str_detect(epsSMASH, "pel") ~ "Pel",
      str_detect(epsSMASH, "alginate") ~ "Alginate",
      str_detect(epsSMASH, "PNAG") ~"PNAG",
    )
  ) %>%
  mutate(compare = paste0(Sp_ID, epsSMASH)) %>%
  select(epsSMASH, compare) %>% 
  full_join(Bundalovic_regions, by = c("compare")) %>%
  mutate(
  Bundalovic = ifelse(Bundalovic == epsSMASH | is.na(epsSMASH), Bundalovic, NA),
  exoPS = ifelse(is.na(Bundalovic), epsSMASH, Bundalovic),
  `exoPS predictions` = case_when(
    is.na(Bundalovic) & !is.na(epsSMASH) ~ "epsSMASH",
    !is.na(Bundalovic) & is.na(epsSMASH) ~ "Bundalovic-Torma et al. (2020)",
    !is.na(Bundalovic) & !is.na(epsSMASH) ~ "Both",
    is.na(Bundalovic) & is.na(epsSMASH) ~ "none"
    )
    )
  

ggplot(epsSMASH_Bundalovic_all_regions %>% group_by(`exoPS predictions`, exoPS) %>% count() %>%
  mutate(`exoPS predictions` = factor(`exoPS predictions`, levels = c("epsSMASH", "Bundalovic-Torma et al. (2020)", "Both"))), 
  aes(x = exoPS, y = n, fill = `exoPS predictions`, label = n)) +
  geom_col(position = position_dodge2(preserve = "single")) +
  scale_y_continuous(trans = "identity") +
  geom_bar_text(position = position_dodge2(preserve = "single"), outside = TRUE, size = 16) +
  ylab("BGCs") +
  scale_fill_brewer(palette =  "Set2") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.title = element_text(size = 16),
    legend.text = element_text(size = 16)
  )

ggsave("figures/all/FigS1.svg", dpi = 300, height = 8, width = 12)


```

```{r}
epsSMASH_regions <- fread("./data/Bundalovic-Torma/all_regions.tsv") %>%
  filter(str_detect(product, 
      "cellulose|pel-positive|pel|alginate|PNAG-ica|PNAG-pga")) %>% 
  filter(product %ni% c("cyanocellulose", "cellulose-gram-positive", "cellulose-gram-positive-like")) %>%
  select(Sp_ID = file, epsSMASH = product) %>%
  mutate(
    epsSMASH = case_when(
      str_detect(epsSMASH, "acetylated-cellulose") ~ "Acetylated-Cellulose",
      str_detect(epsSMASH, "cellulose") ~ "Cellulose",
      str_detect(epsSMASH, "pel") ~ "Pel",
      str_detect(epsSMASH, "alginate") ~ "Alginate",
      str_detect(epsSMASH, "PNAG") ~"PNAG",
    )
  )

Bundalovic_regions <- read_excel("data/Bundalovic-Torma/pcbi.1007721.s009.xlsx", sheet = 1, skip = 1) %>% 
  unite(col = "operon",7:19)

Bundalovic_unique <- Bundalovic_regions %>% filter(Sp_ID %in% setdiff(Bundalovic_regions$Sp_ID, epsSMASH_regions$Sp_ID)) 

epsSMASH_unique <- epsSMASH_regions %>% filter(Sp_ID %in% setdiff(epsSMASH_regions$Sp_ID, Bundalovic_regions$Sp_ID))



gene_functions <- fread("./data/gene_functions.tsv", header = FALSE) %>%
  rename(Query = V1, gene_function = V2)


data <- fread(
  "./data/Bundalovic-Torma/gene_info_Bundalovic.tsv") %>%
  mutate(bitscore = ifelse(is.na(bitscore), 0, bitscore)) %>%
  ungroup() %>%
  mutate(
      region_id = paste0(contig_id, ".", region_id),
      gene_id2 = paste0(genome_id, gene_id),
      location = as.character(location),
      location = gsub(":", "_", location),
      location = gsub("]", "_", location),
      location = gsub("\\[", "", location),
      location = gsub(">", "", location),
      location = gsub("<", "", location),
      location = gsub("join\\{", "", location),
      location = gsub("}", "", location),
      location = gsub(",.*", "", location)
  ) %>%
  group_by(gene_id2) %>%
  separate(`location`, c("start", "end", "strand"), sep = "_") %>%
  mutate(
      start = as.numeric(start),
      end = as.numeric(end),
      strand = ifelse(strand == "(+)", 1, -1)
  ) %>%
  left_join(gene_functions, by = c("Query" = "Query")) %>%
   #filter(!str_detect(product, "-like")) %>%
  filter(str_detect(product, 
      "cellulose|pel-positive|pel|alginate|PNAG-ica|PNAG-pga")) %>% 
  filter(product %ni% c("cyanocellulose", "cellulose-gram-positive", "cellulose-gram-positive-like"))



# Function which plots all the selected gene clusters (every gene cluster in an order for example)
# The function takes a subset size as input, which is the number of gene clusters to plot in each plot
plot_gene_clusters <- function(data, subset_size) {
  # Get the unique regions
  unique_regions <- unique(data$region_id)
  
  # Calculate the number of subsets
  num_subsets <- ceiling(length(unique_regions) / subset_size)
  
  # Loop over each subset
  for(i in seq_len(num_subsets)) {
    # Calculate the indices for this subset
    start_index <- ((i - 1) * subset_size) + 1
    end_index <- min(i * subset_size, length(unique_regions))
    
    # Get the regions for this subset
    subset_regions <- unique_regions[start_index:end_index]
    
    # Filter the data for these regions and plot
    plot <- data %>%
      mutate(strand = ifelse(strand == 1, TRUE, FALSE)) %>%
      filter(region_id %in% subset_regions) %>%
      ggplot(aes(xmin = start, xmax = end, 
      y = product, 
      fill = Query, forward = strand)) +
      geom_gene_arrow() +
      facet_wrap(~ region_id, scales = "free", ncol = 2) +
      ggtitle(paste("Gene clusters from subset", i, "of", num_subsets)) +
      theme(
        axis.text.y = element_text(angle=45)
      )
    print(plot)
  }
}

### Filtering and plotting gene clusters
filter <- data %>%
  filter(genome_id %in% epsSMASH_unique$Sp_ID) %>% filter(Query != "")
  
length(unique(filter$region_id))
plot_gene_clusters(filter, 1)

```


